---
title: 'Lab 10'
author: "Evan Blosser"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```



# Task 1
```{r}
getwd()
```


# Task 2


## Lines A
For line **A** produces a matrix *z* that contains the outer products of *x* and *param*. 
```{r}
# z=outer(x,param,lfun) # A
```


## Lines B
For line **B** sets indices to the maximum of *y*, which is the object made from using *apply(...,sum)* on the object *z*, or the sum of the outer products.
```{r}
# i=max(which(y==max(y))) # B
```


## The Formula for 8 Binomial Experements

$$
\begin{align*}
p(y_1,....,y_8) =& p(y_1)....p(y_8)\\
=& \begin{pmatrix} 20\\y_1 \end{pmatrix} p^{y_1}(1-p)^{20-y_1}\begin{pmatrix} 20\\y_2 \end{pmatrix} p^{y_2}(1-p)^{20-y_2}...\begin{pmatrix} 20\\y_8 \end{pmatrix} p^{y_8}(1-p)^{20-y_8}\\
=& \prod^8_{i=1} \begin{pmatrix} 20\\y_i \end{pmatrix} p^{y_i}(1-p)^{20-y_i}
\end{align*}
$$


## The Code for *mymaxlik()*

```{r}
#max likelihood function  
## For repeated sampling from same distribution                     
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix -- try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun)
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y == max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y)))
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope shoud change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}
```



## Calling the Code & Plot 
The estimate is $0.48$:
```{r}
logbin=function(x,param) log(dbinom(x,prob=param,size=8))
mymaxlik(x=c(3,3,4,3,4,5,5,4),param=seq(0,1,length=1000),lfun=logbin,xlab="x",main="Binomial",cex.main=1 )
```









# Task 3


## *mymaxlik()*
It was found that $\lambda \approx 5.61$ or $\lambda = 5.605606$ from the command line output:
```{r}
logpoiss=function(x,param) log(dpois(x,lambda=param)) 
mymaxlik(x=c(4,6,7,6,5),param=seq(0,20,length=1000),lfun=logpoiss,xlab=expression(lambda),main="Poisson",cex.main=2)
```


## Algebraic Expression
$$
\begin{align*}
p(y) =& (\frac{e^{-\lambda}\lambda 4}{4!}) (\frac{e^{-\lambda}\lambda 6}{6!})(\frac{e^{-\lambda}\lambda 7}{7!})(\frac{e^{-\lambda}\lambda 6}{6!})(\frac{e^{-\lambda}\lambda 5}{5!})\\
=& \prod_{n=4,6,7,6,5}^5 (\frac{e^{-\lambda}\lambda y_i}{y_i!}) 
\end{align*}
$$



## *myNRML()* & Graphical Output
The value for $\hat{\lambda}$ was found to be $5.599999$, which is $-0.005607$ below the calculations made by *mymaxlik()*:
```{r}
T3cmd <- MATH4753EvanB::myNRML(x0=1,delta=0.000001,llik=function(x) log(dpois(4,x)*dpois(6,x)*dpois(7,x)*dpois(6,x)*dpois(5,x)),xrange=c(0,20),parameter="lambda" )
```



## Command Line Output
As can be seen below; as Y approaches approximately zero ,$( 1.776357e-08)$, the x value or solution is approximately $5.599999$ which is the found value for $\hat{\lambda}$ from the function *myNRML()*:
```{r}
T3cmd 
```



# Task 4

## Probability of 2 Heads in 6 Trials

```{r}
logbin=function(x,param) log(dbinom(x,prob=param,size=6))
mymaxlik(x=2,param=seq(0,1,length=1000),lfun=logbin,xlab="x",main="Binomial",cex.main=1 )
```



## Probability of 4 Heads in 10 Trials

```{r}
logbin=function(x,param) log(dbinom(x,prob=param,size=10))
mymaxlik(x=4,param=seq(0,1,length=1000),lfun=logbin,xlab="x",main="Binomial",cex.main=1 )
```





# Task 5


## Algebraic Expression
$$
\begin{align*}
l(\theta_1,\theta_2)  =& bin(y_1│\theta_1 )poiss(y_2 |\theta_2)\\
                      =&  \bigg[ \begin{pmatrix} 20\\y_1 \end{pmatrix} p^{y_1}(1-p)^{20-y_1} \bigg ] \cdot \bigg[ (\frac{e^{-\lambda}\lambda y_2}{y_2!}) \bigg ]
\end{align*}
$$



## Piosson & Binomial 

```{r}
logbinpois=function(theta1,theta2) log(dbinom(4,size=20,prob=theta1)) + log(dpois(4,lambda=theta2))

maxlikg2=function(theta1,theta2,lfun="logbinpois",...){
n1=length(theta1)
n2=length(theta2)
z=outer(theta1,theta2,lfun)
contour(theta1,theta2,exp(z),xlab=expression(theta[1]),ylab=expression(theta[2])) # exp(z) gives the lik
maxl=max(exp(z))    # max lik
coord=which(exp(z)==maxl,arr.ind=TRUE)  # find the co-ords of the max
th1est=theta1[coord[1]] # mxlik estimate of theta1
th2est=theta2[coord[2]]
abline(v=th1est,h=th2est)
axis(3,th1est,round(th1est,2))
axis(4,th2est,round(th2est,2),las=1)
list(th1est=th1est,th2est=th2est)
}
maxlikg2(theta1=seq(0,.5,length=1000),theta2=seq(.5,11.5,length=1000),nlevels=20)
```




# Task 6

## *mymlnorm()*
```{r}
mymlnorm=function(x,mu,sig,...){  #x sample vector
nmu=length(mu) # number of values in mu
nsig=length(sig)
n=length(x) # sample size
zz=c()    ## initialize a new vector
lfun=function(x,m,p) log(dnorm(x,mean=m,sd=p))   # log lik for normal
for(j in 1:nsig){
z=outer(x,mu,lfun,p=sig[j]) # z a matrix 
# col 1 of z contains lfun evaluated at each x with first value of mu, 
# col2 each x with 2nd value of m 
# all with sig=sig[j]
y=apply(z,2,sum)
# y is a vector filled with log lik values, 
# each with a difft mu and all with the same sig[j]
zz=cbind(zz,y)
## zz is the matrix with each column containing log L values, rows difft mu, cols difft sigmas 
}
maxl=max(exp(zz))
coord=which(exp(zz)==maxl,arr.ind=TRUE)
maxlsig=apply(zz,1,max)
contour(mu,sig,exp(zz),las=3,xlab=expression(mu),ylab=expression(sigma),axes=TRUE,
main=expression(paste("L(",mu,",",sigma,")",sep="")),...)
mlx=round(mean(x),2)  # theoretical
mly=round(sqrt((n-1)/n)*sd(x),2)
#axis(1,at=c(0:20,mlx),labels=sort(c(0:20,mlx)))
#axis(2,at=c(0:20,mly),labels=TRUE)
abline(v=mean(x),lwd=2,col="Green")
abline(h=sqrt((n-1)/n)*sd(x),lwd=2,col="Red")

# Now find the estimates from the co-ords
muest=mu[coord[1]]
sigest=sig[coord[2]]

abline(v=muest, h=sigest)
return(list(x=x,coord=coord,maxl=maxl))
}
```



## Graphical Solution
Below is the graphical representation of the solution for $\hat{\mu}$ & $\hat{\sigma}$:
```{r}
mymlnorm(x=c(10,12,13,15,12,11,10),mu=seq(10.4,13.4,length=1000),sig=seq(0.9,3.25,length=1000),lwd=2,labcex=1)
```






# Task 7


## Given Experiment Data
```{r}
sam = rbeta(30,shape1=3,shape2=4)# we know the pop params
```



## *mymlbeta()*
```{r}
mymlbeta=function(x,alpha,beta,...){  #x sample vector
na=length(alpha) # number of values in alpha
nb=length(beta)
n=length(x) # sample size
zz=c()    ## initialize a new vector
lfun=function(x,a,b) log(dbeta(x,shape1=a,shape2=b))   # log lik for beta
for(j in 1:nb){
z=outer(x,alpha,lfun,b=beta[j]) # z a matrix 
# col 1 of z contains lfun evaluated at each x with first value of alpha, 
# col2 each x with 2nd value of a 
# all with b=beta[j]
y=apply(z,2,sum)
# y is a vector filled with log lik values, 
# each with a difft alpha and all with the same sig[j]
zz=cbind(zz,y)
## zz is the matrix with each column containing log L values, rows difft alpha, cols difft betas 
}
maxl=max(exp(zz))    # max lik
coord=which(exp(zz)==maxl,arr.ind=TRUE)  # find the co-ords of the max
aest=alpha[coord[1]] # mxlik estimate of alpha
best=beta[coord[2]]
contour(alpha,beta,exp(zz),las=3,xlab=expression(alpha),ylab=expression(beta),axes=TRUE,
main=expression(paste("L(",alpha,",",beta,")",sep="")),...)

abline(v=aest, h=best)
points(aest,best,pch=19)
axis(4,best,round(best,2),col="Red")
axis(3,aest,round(aest,2),col="Red")
return(list(x=x,coord=coord,maxl=maxl,maxalpha=aest,maxbeta=best))
}
```



## 12 plots
```{r}
layout(matrix(1:12,nr=3,nc=4,byrow=TRUE))
z=c()
a=3
b=4
sam= rbeta(30,shape1=a,shape2=b)
nsam=length(sam)
for(i in 1:12){
w=mymlbeta(x=sample(sam,nsam,replace=TRUE),alpha=seq(0.1,20,length=100),beta=seq(0.1,20,length=100),lwd=2,labcex=1,col="steelblue")
points(a,b,col="Red",pch=19)
abline(v=a,h=b,col="Red")
z<-rbind(z,c(w$maxalpha,w$maxbeta))
 }

```


## Plot Distances
```{r}
colnames(z)=c("alpha","beta")
z
dz=apply(z-c(a,b),1,function(x) sqrt(sum(x^2)) )
plot(dz, ylab ="distances between estimates and known pop. values")
which(dz>4)
```




# Task 8
I included the Newton-Raphson function *myNRML()*:
```{r}
MATH4753EvanB::myNRML(x0=1,delta=0.000001,llik=function(x) log(dpois(12,x)*dpois(10,x)),xrange=c(0,20),parameter="lambda" )
```




