---
title: 'Lab 14'
author: "Evan Blosser"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```


# Task 1
```{r}
getwd()
```



# Task 2

## *mylsq()*
```{r}
mylsq=function(x,y){
ssxx=sum((x-mean(x))^2 )
ssxy=sum((x-mean(x))*(y-mean(y))) ## fill in the missing portion
b1hat=ssxy/ssxx
b0hat=  mean(y)-b1hat*mean(x)   #  fill in the missing portion
return(list(b0hat=b0hat,b1hat= b1hat)) #fill in the missing portion
}

```


## Running 
```{r}
x=1:20
set.seed(29)
y=4+6*x + rnorm(20,0,5) 
T2<- mylsq(x,y)
plot(x,y)
abline(coef = c(T2$b0hat,T2$b1hat),lwd=2,col="blue")

```


## Checking with lm()
As can be seen our coefficients were calculated correctly:
```{r}
slr=lm(y~x)
summary(slr)
```







# Task 3


## *mypreditc()*

```{r}
mypred=function(x,b0,b1){
ym=b0+ b1*x
ym
}
```


## Data from Task 2

### Reload Data
```{r}
x=1:20
set.seed(29)
y=4+6*x + rnorm(20,0,5) 
mylsq(x,y)
```
### Prediction
For $xnew = 15.5$, the prediction was found to be $ym =98.12864$. 
```{r}
xnew<- 15.5
mypred(xnew,3.102214,6.130737)
```
### Plot
```{r}
plot(x,y)
abline(coef = c(T2$b0hat,T2$b1hat),lwd=2,col="blue")+points(x=15.5,y=98.12864, pch = 21, bg = "green")
```


## OJUICE

### A

```{r}
ojuice <- read.csv(file = "OJUICE.csv")
head(ojuice)
T3<-mylsq(x = ojuice$Pectin, y=ojuice$SweetIndex)
plot(ojuice$Pectin,ojuice$SweetIndex,xlab = "Pectin",ylab = "Sweetness Index")
abline(coef = c(T3$b0hat,T3$b1hat),lwd=2,col="blue")
```


### B
The $\hat{\beta_0}$ is the *y* intercept or in this case it it the maximum **sweetness index** and the lowest **pectin** level of the data set. The $\hat{\beta_1}$ is how the **sweetness index** decreases with the increase in  **pectin**.


### C
The **sweetness index** for when the **pectin** in orange juice is $300ppm$ is roughly $5.55888$

```{r}
mypred(x = 300,b0 = T3$b0hat,b1 = T3$b1hat)
```



## DRILLROCK


### A

```{r}
rock <- read.csv(file = "DRILLROCK.csv")
head(rock)
plot(x = rock$DEPTH,y=rock$TIME,xlab = "Drilling Depth",ylab = "Time to Drill 5 ft.")
```



### B
finding coefficients:
```{r}
T3B<-mylsq(x = rock$DEPTH,y=rock$TIME)
T3B
```


Thus the equation is as follows:
$$
\begin{align*}
y = 4.789603 + 0.01438785\cdot x
\end{align*}
$$


### C
```{r}
plot(x = rock$DEPTH,y=rock$TIME,xlab = "Drilling Depth",ylab = "Time to Drill 5 ft.")
abline(coef = c(T3B$b0hat,T3B$b1hat),lwd=2,col="blue")
```



### D
The $\hat{\beta_0}$, or the intercept, Shows the **drilling time** at the shallowest depth. While $\hat{\beta_1}$ shows the increase in **drilling time** as the **drilling depth** increases.


# Task 4

## *mysq()*
```{r}
mysq=function(x,y){
n=length(x)
ssxx=sum((x-mean(x))^2 )
ssxy=sum((x-mean(x))*(y-mean(y))) 
b1hat=ssxy/ssxx
b0hat=  mean(y)-b1hat*mean(x) 
yhat=b0hat+ b1hat*x 
ssr=sum((y-yhat)^2) 
sq= ssr/(n-2)
return(list(ssr=ssr,sq=sq))
}
```


## Carbon


### A

```{r}
carbon <- read.csv("CARBON.csv")
head(carbon)
plot(carbon$LabFurnace,carbon$PilotPlant)
```


### B
```{r}
T4<-mylsq(x = carbon$LabFurnace,y=carbon$PilotPlant)
plot(x = carbon$LabFurnace,y=carbon$PilotPlant,xlab = "lab furnace",ylab = "Pilot Plant")
abline(coef = c(T4$b0hat,T4$b1hat),lwd=2,col="blue")
```



### C
As can be seen below $SSE= 4.694591$ & $S^2=0.2041127$:
```{r}
T4C<-mysq(x= carbon$LabFurnace,y=carbon$PilotPlant )
T4C
```



### D

As can be seen from the calculations below, $s=0.4517883$, which is the standard deviation of the error $\epsilon$, and shows how close the observed *y* values lie compared to the predicted values. In this case the values fall below **1 standard deviation** which shows that the fit and prediction is accurate.   
```{r}
s<-sqrt(T4C$sq)
s
```


