---
title: 'Lab 8'
author: "Evan Blosser"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```



# Task 1
```{r}
getwd()
```
## Note:

See **mylab8.R** for all code comments

# Task 2

## Uniform Distribution
```{r}
uniform <- data.frame(runif(10,0,5))
uniform
```

## Mean and Variance
Given the equations and $\alpha = 0$ and $\beta = 5$:

$$
\begin{align*}
\mu =& \frac{\alpha+\beta }{2} = \frac{0+5}{2} =2.5\\
\mu =& 2.5
\\
\sigma ^2 =&  \frac{ (\beta -\alpha )^2}{12} = \frac{ (5 -0 )^2}{12} = 2.083333\\
\sigma ^2 =& 2.0833
\end{align*}
$$

The mean is $\mu = 2.5$ and the variance is $\sigma ^2 = 2.0833$.

## $\bar{X}$ and $S^2$ using sample:

```{r}
mean(uniform$runif.10..0..5.)
```

$\bar{X}=2.952524$

```{r}
sd(uniform$runif.10..0..5.)^2
```
$S^2=2.832483$

Compared to the population the mean is only off by $0.4$ with the standard deviation has a more significant increase in the sample versus the population.

## sum $T$ & $\bar{Y}$ 
$$
\begin{align*}
E(T) =& nE(Y_i)\\
\\ 
\&
\\
\\
V(\bar{Y}) =& \frac{V(Y_i)}{n}
\end{align*}
$$


## myclt()


```{r}
myclt=function(n,iter){
  y=runif(n*iter,0,5) # A
  data=matrix(y,nr=n,nc=iter,byrow=TRUE) # B
  sm=apply(data,2,sum) # C
  hist(sm)
  sm
}
```


### Line A
Line **A** assigns the object *y* to the output of *runif()* with the settings of $n*iter$ as the number of observations, 0 as the minimum, and 5 as the maximum.

### Line B
Line **B** sets the object *y* to fill the matrix *data* with $n$ rows and $iter$ columns.

### Line C
Line **C** uses *apply()* to use the function *sum* on the matrix *data* and sets it to and object *sm*

### Line D
Line **D** sets the output data of *sm* to an object *w*, that can be printed to or utilized by the user.


## Plot 
```{r}
w=myclt(n=10,iter=10000) # D
```


## Utilizing object *w*
The mean of *w* is $\bar{w} = 24.93929$, and the variance is $S_w^2 = 21.03339$.
```{r}
mean(w)
var(w)
```


## Changing myclt()
```{r}
myclt=function(n,iter){
  y=runif(n*iter,0,5) 
  data=matrix(y,nr=n,nc=iter,byrow=TRUE) 
  ## Changes ################################
  # Apply the mean function to the data matrix
  mean=apply(data,2,mean)
  # Plot a density histogram
  hist(mean, freq = FALSE)
  # Print the mean 
  mean
}
```


## Plotting change myclt() and Utilizing object *w*

The mean of *w* is $\bar{w} = 2.499269$, and the variance is $S_w^2 = 0.2083659$.
```{r}
w=myclt(n=10,iter=10000)
mean(w)
var(w)
```


# Task 3

## mycltu()
```{r}
################### uniform ##########################
### CLT uniform 
## my Central Limit Function
## Notice that I have assigned default values which can be changed when the function is called
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density
ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)
}
```


##	w=apply(data,2,mean), how does the apply function use the 2?
The *apply()* function uses the *2* as a MARGIN, indicating we want to select columns as the target of *mean*.

## How many terms are in w, when mycltu(n=20,iter=100000) is called?
there are $101$ X & Y terms.


## curve(dnorm(x,mean=(a+b)/2,
This defines a theoretical curve using *dnorm()*, meaning that it is **Normal**, and the mean is as defined in **Task 2** for a uniform distribution.


## sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3): 
This is the standard deviation for the *dnorm()* curve as in **Task 2**; however, as that was the variance, and the standard deviation is needed th formula takes on the following form for $n$ sample size:

$$
\begin{align*}
\frac{b-a}{\sqrt{12*n}}
\end{align*}
$$

$$
\begin{align*}
\hspace{2.5mm} NOT \hspace{2.5mm} \frac{(b-a)^2}{12}
\end{align*}
$$


## Plots for $n= 1,2,3,5,10,30$ 
```{r}
w = mycltu(n=1,iter=10000,a = 0,b = 10)
w = mycltu(n=2,iter=10000,a = 0,b = 10)
w = mycltu(n=3,iter=10000,a = 0,b = 10)
w = mycltu(n=5,iter=10000,a = 0,b = 10)
w = mycltu(n=10,iter=10000,a = 0,b = 10)
w = mycltu(n=30,iter=10000,a = 0,b = 10)
```

## Conclusion
After examining the plots above it would seem like a size of $n=5$ gives the most even or *uniform* distribution considering a bell curve. When $n=1$ the values all have an identical mean, and for $n=30$ the mean is centered around 5 with a propability density above $0.6$. 



# Task 4

## mycltb()
```{r}
##############################  Binomial #########
## CLT Binomial
## CLT will work with discrete or continuous distributions 
## my Central Limit Function
## Notice that I have assigned default values which can be changed when the function is called
mycltb=function(n,iter,p=0.5,...){
  ## r-random sample from the Binomial
  y=rbinom(n*iter,size=n,prob=p)
  ## Place these numbers into a matrix
  ## The columns will correspond to the iteration and the rows will equal the sample size n
  data=matrix(y,nr=n,nc=iter,byrow=TRUE)
  ## apply the function mean to the columns (2) of the matrix
  ## these are placed in a vector w
  w=apply(data,2,mean)
  ## We will make a histogram of the values in w
  ## How high should we make y axis?
  ## All the values used to make a histogram are placed in param (nothing is plotted yet)
  param=hist(w,plot=FALSE)
  ## Since the histogram will be a density plot we will find the max density
  
  ymax=max(param$density)
  ## To be on the safe side we will add 10% more to this
  ymax=1.1*ymax
  
  ## Now we can make the histogram
  ## freq=FALSE means take a density
  hist(w,freq=FALSE,  ylim=c(0,ymax),
       main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
       xlab="Sample mean",...)
  ## add a density curve made from the sample distribution
  #lines(density(w),col="Blue",lwd=3) # add a density plot
  ## Add a theoretical normal curve 
  curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 
  
}
```


## Plots for $n = 4,5,10,20$ 
```{r}
mycltb(n=4,iter=10000,p=0.3)
mycltb(n=5,iter=10000,p=0.3)
mycltb(n=10,iter=10000,p=0.3)
mycltb(n=20,iter=10000,p=0.3)
```




## Plots for $n = 4,5,10,20$ & $p=0.7$
```{r}
mycltb(n=4,iter=10000,p=0.7)
mycltb(n=5,iter=10000,p=0.7)
mycltb(n=10,iter=10000,p=0.7)
mycltb(n=20,iter=10000,p=0.7)
```


## Plots for $n = 4,5,10,20$ & $p=0.5$
```{r}
mycltb(n=4,iter=10000,p=0.5)
mycltb(n=5,iter=10000,p=0.5)
mycltb(n=10,iter=10000,p=0.5)
mycltb(n=20,iter=10000,p=0.5)
```


## Conclusion
The higher the sample size $n$ is the more uniform the distribution is, while the lower the number actually yields gaps such as in the graphs of $n=4$. The probability also has more of an effect on the distribution when the sample size is low. This can be seen in the graph for $n=4$ & $p=0.3$, the distribution skews to the left while the opposite occurs for $p=0.7$. The probability has little to no significant affect on the sample size $n=20$; thus, it can be concluded that the probability affects the binomial distribution for samples $n<20$.








# Task 5



## mycltp()
```{r}
####### Poisson ######################
## CLT Poisson
## CLT will work with discrete or continuous distributions 
## my Central Limit Function
## Notice that I have assigned default values which can be changed when the function is called
mycltp=function(n,iter,lambda=10,...){
  ## r-random sample from the Poisson
  y=rpois(n*iter,lambda=lambda)
  ## Place these numbers into a matrix
  ## The columns will correspond to the iteration and the rows will equal the sample size n
  data=matrix(y,nr=n,nc=iter,byrow=TRUE)
  ## apply the function mean to the columns (2) of the matrix
  ## these are placed in a vector w
  w=apply(data,2,mean)
  ## We will make a histogram of the values in w
  ## How high should we make y axis?
  ## All the values used to make a histogram are placed in param (nothing is plotted yet)
  param=hist(w,plot=FALSE)
  ## Since the histogram will be a density plot we will find the max density
  ymax=max(param$density)
  ## To be on the safe side we will add 10% more to this
  ymax=1.1*ymax
  ## Make a suitable layout for graphing
  layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))
  ## Now we can make the histogram
  hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
       main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
       xlab="Sample mean",...)
  ## add a density curve made from the sample distribution
  #lines(density(w),col="Blue",lwd=3) # add a density plot
  ## Add a theoretical normal curve 
  curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  # Now make a new plot
  # Since y is discrete we should use a barplot
  barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
  x=0:max(y)
  plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
       main="Probability function for Poisson", ylab="Probability",xlab="y")
}
```



## Plots for $n = 2,3,5,10,20$
```{r}
mycltp(n=2, iter=10000,lambda=4)
mycltp(n=3, iter=10000,lambda=4)
mycltp(n=5, iter=10000,lambda=4)
mycltp(n=10, iter=10000,lambda=4)
mycltp(n=20, iter=10000,lambda=4)
```


## Plots for $n = 2,3,5,10,20$ & $\lambda=10$
```{r}
mycltp(n=2, iter=10000,lambda=10)
mycltp(n=3, iter=10000,lambda=10)
mycltp(n=5, iter=10000,lambda=10)
mycltp(n=10, iter=10000,lambda=10)
mycltp(n=20, iter=10000,lambda=10)
```

## Video
<video width="320" height="240" controls>
  <source src="vid.mp4" type="video/mp4">
Your browser does not support the video tag.
</video>





# Task 6

I chose the *mycltp()* function and created it using **Roxygen** for proper documentation. Below is an example of it running from my package *MATH4753EvanB*.

```{r}
MATH4753EvanB::mycltp(n=2, iter=10000,lambda=4)
```


