---
title: "Lab1_Evan_A_Blosser"
author: "Evan Blosser"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
ddt <- read.csv(file = "DDT.csv")
head(ddt)
```

## Task 3

The qualitative variables are **RIVER** and **SPECIES**

The quantitative variables are **MILE**, **LENGTH**, **WEIGHT**, and **DDT**

There are 3 Types of **SPECIES**: LMBASS, CCATFISH, SMBUFFALO with a total of 144
```{r}
# Set SpeciesTab as the Species Table 
SpeciesTab <- table(ddt$SPECIES)     
barplot(SpeciesTab, col = rainbow(3))      
```
```{r}
# Call for Data Summary 
summary(ddt)            
```


```{r}
SpeciesSubLMB=with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
SpeciesSubLMB
```
```{r}
SpeciesSubRiver=with(ddt, ddt[DDT>4.0 & RIVER=="SCM",])
SpeciesSubRiver
```



### Clicker Questions

### Question 1
From the Summary in Task 3, the mean **LENGTH** of fish is 42.81
 
### Question 2
```{r}
Weight <- ddt$WEIGHT
sd(Weight)
```
The Standard deviation of the weight of fish within the ddt data frame is:
  376.5461 grams
  
### Question 3
This is not the correct graph as the Y & X axis are labled wrong as can be seen from the correct graph below
```{r}
#Use the data option
plot(WEIGHT~LENGTH,data=ddt)
```

### Question 4 
As can be seen below the last value of v/20 is **1.00**
```{r}
v = 1:20
v/20
```




## Task 4
```{r}
# Set SpeciesTab as the Species Table 
RiversTab=with(ddt,table(RIVER))     
RiversTab                           
# Create Barplot                   
barplot(RiversTab, col = rainbow(4))     
# Set table of Rivers Crossed w/ Species  
RiverXFish=with(ddt,table(RIVER,SPECIES)) 
RiverXFish
# Create Barplot 
barplot(RiverXFish, col = rainbow(4))
```

## Task 5
```{r}
pie(SpeciesTab,col=rainbow(4))
pie(RiversTab,col=rainbow(4))
```

## Task 6
```{r}
layout(matrix(c(1,2,3),nr=1,nc=3))# 1 row 3 cols
with(ddt,boxplot(LENGTH,ylab="LENGTH",col="Blue",notch=TRUE))
with(ddt,boxplot(WEIGHT,ylab="WEIGHT",col="Green",notch=TRUE))
with(ddt,boxplot(MILE,ylab="MILE",col="Red",notch=TRUE))
```

## Task 7
```{r}
#
coplot(WEIGHT~LENGTH|RIVER,data=ddt,col= rainbow(4))
#
coplot(DDT~WEIGHT|SPECIES,data=ddt,col=rainbow(4))
```

## Task 8

```{r}
#Using dplyr
library(dplyr)
library(ggplot2)
# notice how the infix %>% operator works ( the pipe )

ddt %>% ggplot(aes(x=SPECIES,y=WEIGHT,fill=RIVER)) + geom_boxplot() + ggtitle("Evan Blosser")
```
```{r}
ddt %>% ggplot(aes(x=RIVER,y=LENGTH,fill=SPECIES)) + geom_violin() + ggtitle("Evan Blosser")
```

```{r}
ggplot(ddt, aes(x=WEIGHT, y=LENGTH)) + geom_point(aes(color = SPECIES)) + ggtitle("Evan Blosser")

```




