---
title: "Lab 2"
author: "Evan Blosser"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Task 1
```{r}
getwd()
```


# Task 2
```{r}
EPAGAS <- read.csv(file = "EPAGAS.csv")
head(EPAGAS)
```


# Task 3

## Bullet Point 1 & 2
```{r}
mpg <- EPAGAS$MPG
mpg
z <- (mpg-mean(mpg))/sd(mpg)
```
As shown below the mean of **z** was calculated to be approximately 0 (1.058709e-16); while the square root of the standard deviation of **z** was calculated to be 1:
```{r}
mean(z)
sd(z)^2
```

## Bullet Point 3 & 4
The outlier of **MPG** was found to be $44.9$.

While the possible outliers are $30.0$, $42.1$, and $31.8$.

```{r}
# Find the z values greater than 3 in size
z[abs(z)>3]

# Find the I values corresponding to these z values
mpg[abs(z)>3]

# Find the values of z that are possible outliers
z[abs(z)>=2 & abs(z)<=3]

#Find the values of I which are possible outliers
mpg[abs(z)>=2 & abs(z)<=3]
``` 

## Bullet Point 5
 
```{r}
# Plot the outliers in red and the possible outliers in blue
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))  

#Plot points with different plotting shapes        
mypch = ifelse(abs(z)>3, 20,
        ifelse(abs(z)>=2 &abs(z)<=3,21, 22))
library(lattice)
dotplot(mpg,col=mycol)
```


# Task 4

## Bullet Point 1
```{r}
boxplot(mpg,main="Miles Per Gallon: Task 4 Outlier identification",ylab="mpg",col="Blue",notch=TRUE)
```

## Bullet Point 2

Chebyshev's theorem states that generally at least: 
$$
 \big( 1- \frac{1}{k^2} \big)    
$$
measruments will fall within k standard deviations of the mean. Or the intervals fopr samples and populations respectively.
$$
(\bar{y} \pm ks)
$$
$$
 (\sigma \pm ks) 
$$
Given that we need to predict the proportion of data within 2 standard deviations of the mpg mean:
$$
 \big( 1- \frac{1}{2^2} \big) = \frac{3}{4}   
$$
it can be seen that the proportion of data within 2 standard deviations of the mpg mean is $\frac{3}{4}$ of the data using **Chebyshev's theorem**.

## Bullet Point 3

Using **R** the exact proportion was calculated to be **0.96**.
```{r}
# Actual Proportion
length(mpg[abs(z)<2])/length(mpg)
```
## Bullet Point 4
**Chebyshev's theorem** does agree with the data in that there is at least $\frac{3}{4}$ of the data. However, this is fairly inaccurate as the actual proportion of data within 2 standard deviations of the mean was **0.96** or $96\%$.

## Bullet Point 5

By the **Empirical Rule** the proportion of data is $95\%$ for 2 standard deviations of he mean.

## Bullet Point 6

I would say that the **Empirical Rule** corresponds more closely to the calculated proportion as compared to **Chebyshev's theorem**; as the **Empirical Rule** was off by only $1\%$.

## Bullet Point 7

I would say that the **Empirical Rule** is more valid in this case, becasue it accuratly described the proportion of data  within 2 standard deviations of the mean as compared to **Chebyshev's theorem**. 
