---
title: 'Lab6'
author: "Evan Blosser"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---
```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```


## Task 1
```{r}
getwd()
```


## Task 2 {.tabset}

### 4 Normal Curves

```{r}
# Sets Layout for 4 graphs 2 on top 2 on bottom
layout(matrix(1:4, nr=2,nc=2))
# Normal Curve Plots
curve(dnorm(x, mean=10,sd=4),xlim=c(-6,25),ylab = "sd=4")
curve(dnorm(x, mean=10,sd=2),xlim=c(2,17),ylab = "sd=2")
curve(dnorm(x, mean=5,sd=10),xlim=c(-30,40),ylab = "sd=10")
curve(dnorm(x, mean=5,sd=0.5),xlim=c(3,7),ylab = "sd=0.5")
```



### Normal mean = 0 & sd = 1

```{r}
curve(dnorm(x, mean=0,sd=1),xlim=c(-3,3),ylab = "sd=1")
# In order to find the area between x=10 and 25
#  - seq defines desired x points
xcurve=seq(2,5,length=1000)
# For The Y part of the curve  we use the X points
#  and the data for the curve in a normal distribution
ycurve=dnorm(xcurve,mean=0,sd=1)
# Creates a Polygon and fills it bounded by the given
#   x & y curves
polygon(c(2,xcurve,5),c(0,ycurve,0),col="Red")
# Put in the text with the appropriate area
# Area
area=pnorm(5,mean=0,sd=1)-pnorm(2,mean=0,sd=1)
area=round(area,4)
# Click to paste the text onto the graph
text(x = 0, y = 0.2, paste("Area = ", area, sep=""))
```



###  Normal mean = 4 & sd = 2

```{r}
curve(dnorm(x, mean=4,sd=2),xlim=c(-10,10),ylab = "sd=2")
# In order to find the area between x=10 and 25
#  - seq defines desired x points
xcurve=seq(1,5,length=1000)
# For The Y part of the curve  we use the X points
#  and the data for the curve in a normal distribution
ycurve=dnorm(xcurve,mean=4,sd=2)
# Creates a Polygon and fills it bounded by the given
#   x & y curves
polygon(c(1,xcurve,5),c(0,ycurve,0),col="green")
# Put in the text with the appropriate area
# Area
prob=pnorm(5,mean=4,sd=2)-pnorm(1,mean=4,sd=2)
prob=round(prob,4)
# Click to paste the text onto the graph
text(x = 0, y = 0.2, paste("Area = ", prob, sep=""))
```




###  Normal mean = 10 & sd = 4
```{r}
curve(dnorm(x, mean=10,sd=4),xlim=c(-10,20),ylab = "sd=4")
#
# In order to find the area between x=10 and 25
#  - seq defines desired x points
xcurve=seq(-10,10,length=1000)
# For The Y part of the curve  we use the X points
#  and the data for the curve in a normal distribution
ycurve=dnorm(xcurve,mean=10,sd=4)
# Creates a Polygon and fills it bounded by the given
#   x & y curves
polygon(c(-10,xcurve,10),c(0,ycurve,0),col="blue")
# Put in the text with the appropriate area
# Area
prob=pnorm(10,mean=10,sd=4)
prob=round(prob,4)

# Click to paste the text onto the graph
text(x = 0, y = 0.09, paste("Area = ", prob, sep=""))
```







###  Normal mean = -2 & sd = 1/2
```{r}
curve(dnorm(x, mean=-2,sd=0.5),xlim=c(-4,1),ylab = "sd=0.5")
# In order to find the area between x=10 and 25
#  - seq defines desired x points
xcurve=seq(-3,-2,length=1000)
# For The Y part of the curve  we use the X points
#  and the data for the curve in a normal distribution
ycurve=dnorm(xcurve,mean=-2,sd=0.5)
# Creates a Polygon and fills it bounded by the given
#   x & y curves
polygon(c(-3,xcurve,-2),c(0,ycurve,0),col="purple")
# Put in the text with the appropriate area
# Area
prob=pnorm(-2,mean=-2,sd=0.5)-pnorm(-3,mean=-2,sd=0.5)
prob=round(prob,4)
# Click to paste the text onto the graph
text(x = 0, y = 0.2, paste("Area = ", prob, sep=""))
```






## Task 3{.tabset}
The following graphs were made with the script the commented version of lab6.R:


### Overlayed Gamma Distribution set

```{r}
curve(dgamma(x,shape=1,scale=1),xlim=c(0,10),ylim=c(0,1),col="Red",lwd=2,
ylab="Gamma density", main="Beta=1")
# Text Click Paste
text(x= 1.5,y=0.5 ,paste("alpha=",1))
# Curve based of Gamma Distribution
curve(dgamma(x,shape=3,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE,lwd=2)
# Text Click Paste
text(x= 2,y= 0.32,paste("alpha=",3))
# Curve based of Gamma Distribution
curve(dgamma(x,shape=5,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE, col="Blue",lwd=2)
# Text Click Paste
text(x= 4,y= 0.25,paste("alpha=",5))
```


### Gamma Distribution: shape=3, scale=2

```{r}
curve(dgamma(x,shape=3,scale=2),xlim=c(0,10),ylim=c(0,0.3),col="blue",lwd=2,ylab="Gamma density", main="Gamma Distribution: shape=3, scale=2")
xgamma1 = seq(2,5,length=1000)
ygamma1 = dgamma(xgamma1,shape = 3,scale = 2)
polygon(c(2,xgamma1,5),c(0,ygamma1,0),col = "Purple")
gammaprob1 = pgamma(5,shape = 3,scale = 2) - pgamma(2,shape = 3,scale = 2)
gammaprob1 = round(gammaprob1,4)
text(x= 4,y= 0.25,paste("P(2<Y<5)=",gammaprob1,sep = ""))
```

### Gamma Distribution: shape=6, scale=3 

```{r}

curve(dgamma(x,shape=6,scale=3),xlim=c(1,30),ylim=c(0,.06),col="blue",lwd=2,ylab="Gamma density", main="Gamma Distribution: shape=6, scale=3 ")

xgamma2 = seq(1,4,length=1000)
ygamma2 = dgamma(xgamma2,shape = 6,scale = 3)
polygon(c(1,xgamma2,4),c(0,ygamma2,0),col = "Purple")
gammaprob2 = pgamma(4,shape = 6,scale = 3) - pgamma(1,shape = 6,scale = 3)
gammaprob2 = round(gammaprob2,4)

text(x= 15,y= 0.04,paste("P=",gammaprob2,sep = ""))

```


### Gamma Distribution: shape=2, scale=4

```{r}
curve(dgamma(x,shape=2,scale=4),xlim=c(0,30),ylim=c(0,0.1),col="blue",lwd=2,ylab="Gamma density", main="Gamma Distribution: shape=2, scale=4 ")

xgamma3 = seq(3,6,length=1000)
ygamma3 = dgamma(xgamma3,shape = 2,scale = 4)
polygon(c(3,xgamma3,6),c(0,ygamma3,0),col = "Purple")
gammaprob3 = pgamma(6,shape = 2,scale = 4) - pgamma(3,shape = 2,scale = 4)
gammaprob3 = round(gammaprob3,4)

text(x= 15,y= 0.05,paste("P=",gammaprob3,sep = ""))
```




## Task 4{.tabset}

### 4 plots in layout()
```{r}
# Chi -sq ### alpha = df/2, beta=2
layout(matrix(1:4, nr=2,nc=2))
# Curve Based off Chi-Square distribution
curve(dchisq(x,df=1),xlim=c(0,8),ylim=c(0,1),col="Red",lwd=2,ylab="Chisq density", main="df=1")
curve(dchisq(x,df=2),xlim=c(0,10),ylim=c(0,1),col="Blue",lwd=2,ylab="Chisq density",main="df=2")
curve(dchisq(x,df=4),xlim=c(0,10),ylim=c(0,.3),col="Green",lwd=2,ylab="Chisq density", main="df=4")
curve(dchisq(x,df=20),xlim=c(0,40),ylim=c(0,.1),col="Purple",lwd=2,ylab="Chisq density",main="df=20")
```

### Chi-Square: Degree of Freedom = 2
```{r}
curve(dchisq(x,df=2),xlim=c(0,10),ylim=c(0,1),col="Blue",lwd=2,ylab="Chisq density",main="df=2")

xchi1 = seq(2,4,length=1000)
ychi1 = dchisq(xchi1,df=2)
polygon(c(2,xchi1,4) ,c(0,ychi1,0),col = "purple")

chiprob1 = pchisq(4,df = 2) - pchisq(2,df = 2)
chiprob1 = round(chiprob1,4)

text(x= 5,y= 0.25,paste("Area=",chiprob1,sep = ""))
```


### Chi-Square: Degree of Freedom = 3
```{r}
curve(dchisq(x,df=3),xlim=c(0,10),ylim=c(0,0.27),col="forestgreen",lwd=2,ylab="Chisq density", main="df=3")

xchi2 = seq(3,5,length=1000)
ychi2 = dchisq(xchi2,df=3)
polygon(c(3,xchi2,5) ,c(0,ychi2,0),col = "cyan")

chiprob2 = pchisq(5,df = 3) - pchisq(3,df = 3)
chiprob2 = round(chiprob2,4)

text(x= 3,y= 0.25,paste("Area=",chiprob2,sep = ""))

```


### Chi-Square: Degree of Freedom = 20
```{r}
curve(dchisq(x,df=20),xlim=c(0,40),ylim=c(0,.08),col="Purple",lwd=2,ylab="Chisq density",main="df=20")

xchi3 = seq(10,21,length=1000)
ychi3 = dchisq(xchi3,df=20)
polygon(c(10,xchi3,21) ,c(0,ychi3,0),col = "cyan")

chiprob3 = pchisq(21,df = 20) - pchisq(10,df = 20)
chiprob3 = round(chiprob3,4)

text(x= 17,y= 0.025,paste("Area=",chiprob3,sep = ""))

```


## Task 5{.tabset} 


### *4 Weibull plots*
```{r}
layout(matrix(1:4, nr=2,nc=2))
curve(dweibull(x,shape = 5,scale = 2),xlim=c(0,4),ylim=c(0,1),col="Purple",lwd=2,ylab="Weibull density",main="shape = 5, scale = 2")

curve(dweibull(x,shape = 10,scale = 3.5),xlim=c(1,6),ylim=c(0,1.5),col="forestgreen",lwd=2,ylab="Weibull density",main="shape = 10,scale = 3.5")

curve(dweibull(x,shape = 20,scale = 4),xlim=c(1,6),ylim=c(0,2),col="blue",lwd=2,ylab="Weibull density",main="shape = 20,scale = 4")

curve(dweibull(x,shape = 8,scale = 4),xlim=c(0,6),ylim=c(0,1),col="red",lwd=2,ylab="Weibull density",main="shape = 8, scale = 4")
```


###  $shape = 5, scale = 2:  P(1\leq Y\leq2)$
```{r}
curve(dweibull(x,shape = 5,scale = 2),xlim=c(0,6),ylim=c(0,1),col="Purple",lwd=2,ylab="Weibull density",main="shape = 5, scale = 2")

xWei1 = seq(1,2,length=1000)
yWei1 = dweibull(xWei1,shape = 5,scale = 2)

polygon(c(1,xWei1,2) ,c(0,yWei1,0),col = "cyan")

Weiprob1 = pweibull(2,shape = 5,scale = 2) - pweibull(1,shape = 5,scale = 2)
Weiprob1 = round(Weiprob1,4)

text(x= 4,y= 0.4,paste("Area=",Weiprob1,sep = ""))

```


###  $shape = 10, scale = 3.5: P(2\leq Y\leq3)$
```{r}
curve(dweibull(x,shape = 10,scale = 3.5),xlim=c(1,6),ylim=c(0,1.5),col="green",lwd=2,ylab="Weibull density",main="shape = 10,scale = 3.5")

xWei2 = seq(2,3,length=1000)
yWei2 = dweibull(xWei2,shape = 10,scale = 3.5)

polygon(c(2,xWei2,3) ,c(0,yWei2,0),col = "cyan")

Weiprob2 = pweibull(3,shape = 10,scale = 3.5) - pweibull(2,shape = 10,scale = 3.5)
Weiprob2 = round(Weiprob2,4)

text(x= 5,y= 0.4,paste("Area=",Weiprob2,sep = ""))

```



###  $shape = 20,scale = 4: P(3\leq Y\leq 4)$
```{r}
curve(dweibull(x,shape = 20,scale = 4),xlim=c(0,6),ylim=c(0,2),col="blue",lwd=2,ylab="Weibull density",main="shape = 20,scale = 4")

xWei3 = seq(3,4,length=1000)
yWei3 = dweibull(xWei3,shape = 20,scale = 4)

polygon(c(3,xWei3,4) ,c(0,yWei3,0),col = "cyan")

Weiprob3 = pweibull(4,shape = 20,scale = 4) - pweibull(3,shape = 20,scale = 4)
Weiprob3 = round(Weiprob3,4)

text(x= 2,y= 0.4,paste("Area=",Weiprob3,sep = ""))

```



###  $shape = 8, scale = 4: P(3\leq Y\leq 4)$
```{r}
curve(dweibull(x,shape = 8,scale = 4),xlim=c(0,10),ylim=c(0,1),col="red",lwd=2,ylab="Weibull density",main="shape = 8, scale = 4")


xWei4 = seq(3,4,length=1000)
yWei4 = dweibull(xWei4,shape = 8,scale = 4)

polygon(c(3,xWei4,4) ,c(0,yWei4,0),col = "cyan")

Weiprob4 = pweibull(4,shape = 8,scale = 4) - pweibull(3,shape = 8,scale = 4)
Weiprob4 = round(Weiprob4,4)

text(x= 6,y= 0.4,paste("Area=",Weiprob4,sep = ""))

```



## Task 6 {.tabset} 

### *4 Beta Plots*
```{r}
layout(matrix(1:4, nr=2,nc=2))
curve(dbeta(x,shape1 = 5,shape2 =2),xlim=c(-1,2),ylim=c(0,4),col="Purple",lwd=2,ylab="Beta density",main="shape1 = 5,shape2 =2")

curve(dbeta(x,shape1 = 6,shape2 = 3),xlim=c(-1,2),ylim=c(0,3),col="forestgreen",lwd=2,ylab="Beta density",main="shape1 = 6,shape2 = 3")

curve(dbeta(x,shape1 = 3,shape2 =4),xlim=c(-1,2),ylim=c(0,2.5),col="blue",lwd=2,ylab="Beta density",main="shape1 = 3,shape2 =4")

curve(dbeta(x,shape1 = 2,shape2 = 5),xlim=c(-1,2),ylim=c(0,3),col="red",lwd=2,ylab="Beta density",main="shape1 = 2,shape2 = 5")
```

### $shape1 = 5, shape2 = 2: P(0.5\leq Y\leq 1)$

```{r}
curve(dbeta(x,shape1 = 5,shape2 =2),xlim=c(0,2),ylim=c(0,4),col="Purple",lwd=2,ylab="Beta density",main="shape1 = 5,shape2 =2")

xB1 = seq(.5,1,length=1000)
yB1 = dbeta(xB1,shape1 = 5,shape2 =2)

polygon(c(.5,xB1,1) ,c(0,yB1,0),col = "cyan")

Bprob1 = pbeta(1,shape1 = .5,shape2 =.2) - pbeta(.5,shape1 = .5,shape2 =.2)
Bprob1 = round(Bprob1,4)

text(x= 1.5 ,y= 0.4,paste("Area=",Bprob1,sep = ""))

```



### $shape1 = 6,shape2 = 3: P(.2\leq Y\leq .5)$

```{r}
curve(dbeta(x,shape1 = 6,shape2 = 3),xlim=c(0,1),ylim=c(0,3),col="forestgreen",lwd=2,ylab="Beta density",main="shape1 = 6,shape2 = 3")

xB1 = seq(.2,.5,length=1000)
yB1 = dbeta(xB1,shape1 = 6,shape2 = 3)

polygon(c(.2,xB1,.5) ,c(0,yB1,0),col = "cyan")

Bprob1 = pbeta(.5,shape1 = 6,shape2 = 3) - pbeta(.2,shape1 = 6,shape2 = 3)
Bprob1 = round(Bprob1,4)

text(x= .75,y= 0.4,paste("Area=",Bprob1,sep = ""))

```



### $shape1 = 3,shape2 =4: P( Y\leq .5)$

```{r}
curve(dbeta(x,shape1 = 3,shape2 =4),xlim=c(-.5,1),ylim=c(0,2.5),col="blue",lwd=2,ylab="Beta density",main="shape1 = 3,shape2 =4")

xB1 = seq(0,.5,length=1000)
yB1 = dbeta(xB1,shape1 = 3,shape2 =4)

polygon(c(0,xB1,.5) ,c(0,yB1,0),col = "cyan")

Bprob1 = pbeta(.5,shape1 = 3,shape2 =4)
Bprob1 = round(Bprob1,4)

text(x= -.4,y= 0.4,paste("Area=",Bprob1,sep = ""))

```




### $shape1 = 2,shape2 = 5: P(.25\leq Y\leq .75)$

```{r}
curve(dbeta(x,shape1 = 2,shape2 = 5),xlim=c(-.2,1),ylim=c(0,3),col="red",lwd=2,ylab="Beta density",main="shape1 = 2,shape2 = 5")

xB1 = seq(.25,.75,length=1000)
yB1 = dbeta(xB1,shape1 = 2,shape2 = 5)

polygon(c(.25,xB1,.75) ,c(0,yB1,0),col = "cyan")

Bprob1 = pbeta(.75,shape1 = 2,shape2 = 5) - pbeta(.25,shape1 = 2,shape2 = 5)
Bprob1 = round(Bprob1,4)

text(x= .8,y= 0.4,paste("Area=",Bprob1,sep = ""))

```











## Task 7{.tabset} 


### myncurve() running 
```{r}
MATH4753EvanB::myncurve(10,4,10)
```




###  Comparing to Previous Task 2: mean = 10 & sd = 4
```{r}
curve(dnorm(x, mean=10,sd=4),xlim=c(-10,20),ylab = "sd=4")
#
# In order to find the area between x=10 and 25
#  - seq defines desired x points
xcurve=seq(-10,10,length=1000)
# For The Y part of the curve  we use the X points
#  and the data for the curve in a normal distribution
ycurve=dnorm(xcurve,mean=10,sd=4)
# Creates a Polygon and fills it bounded by the given
#   x & y curves
polygon(c(-10,xcurve,10),c(0,ycurve,0),col="blue")
# Put in the text with the appropriate area
# Area
prob=pnorm(10,mean=10,sd=4)
prob=round(prob,4)

# Click to paste the text onto the graph
text(x = 0, y = 0.09, paste("Area = ", prob, sep=""))
```

